Viewing Computer Science through Citation Analysis

### 2020-05-11 Scientometrics/rebuttal2/rebuttal_letter.tex

\item Section-wise brief discussion is missing at the end of Introduction.

\emph{We have revised the Introduction to better track the flow of information through the document}

\item Quality of some images is really poor. Please improve them to make it much more visible.

\emph {We are a little puzzled by this comment, since the pdfs generated on our machines show acceptable image quality. However, we have regenerated all the images in pdf (and TIFF @600 dpi as supplementary images for the journal). In some cases, we also modified axis labels to make them more readable. Thank you for flagging this defect. In the event that the manuscript is accepted for publication, we will certainly work closely with journal staff to generate even higher resolution figures.}

@@

\emph{A correction has been made}

\item C04. For the section ``1 Introduction," what is the purpose of this research, or what is this research aiming at? Ultimately, it is estimated that the part "we extend the work of Salton and Bergmark" will be the purpose. 
However, readers cannot understand the originality or significance of this research, because it seems to simply apply an existing methodology to current data. Thus, you should describe what you want to know from your 
results using the work of Salton and Bergmark. First, you should make the purpose specific in the section "1 Introduction". Second, answer to your new purpose in the section ``3 Discussion" or others (i.g. "Conclusion" if you add).
I recommend that you write an outspoken expression using "purpose" or "aim at" about your purpose for appealing your research's value to roughly-looking readers.

\emph{We interpret this comment from the reviewer as noting a lack of communication of purpose in the manuscript. We agree that in trying to reach out to multiple readerships with sometimes thinly overlapping interests, we have 
presented a broad sense of purpose. Therefore, we have tried to clarify purpose in the revision.}

\emph{Regarding `existing methodology to current data.', we would say that the purpose of this paper was discovery. However, we will also offer that while we use Graclus, an existing methodology, for clustering by direct citation, it has not been used exhaustively in citation analysis as far as we can tell, and this is consistent with text in \v{S}ubelj, van Eck, and Waltman (2016). Further, we have modified co-citation based clustering as described by Small and Sweeney (1985). Lastly, that clustering methods tend to display sensitivity to input data such that every case has to be carefully evaluated. However, it is our responsibility to communicate all this to the reader, so the reviewer's point is very well taken and we thank her/him for it}

@@

\item C06. [Optional] In the section ``3 Discussion,"  you had better to write a limitation by excluding small fields, which may include innovative seeds. In order to use this research for the purpose of promoting academic-industry collaboration, you can also describe it as your future work. I like Fig.3 (c), because I'm excited to low score areas that may be new opportunities. Meanwhile, currently, agricultural IT is emerging and fishery starts to use IT. You may be able to provide information of new opportunities with your cutoff small fields by using figures like Fig.3 (c)

\emph{For reasons of scope and feasibility, we restricted ourselves to interpreting strong signals from a measurement at a single point in time. Further, `To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster', but we have added a sentence to the Discussion about excluding small fields, which may (or may not) include innovative seeds}

\item C07. [Optional] As a significance of your research, the authors can suggest the academic about a question if some current systems that a paper connects to one field are outdated.

\emph{Even though we find the idea attractive, we are hesitant to verge into excessive speculation. We do make the point that contextualized interpretation of these results is important and will address the suggestion to a greater extent in the revision}

### 2020-05-11 Scientometrics/rebuttal2/comp_revised_2.tex
R2

@@

\begin{abstract}
Computer science has experienced dramatic growth and diversification over the last twenty years. Towards a current understanding of the structure of this discipline, we analyze 
a \replaced{large sample}{cohort} of the computer science literature from the DBLP database. For insight on the features of this cohort and the relationship within its components, we have constructed article 
level clusters based on either direct citations or co-citations, and reconciled them \replaced{with}{to} major and minor subject categories in the All Science Journal Classification (ASJC). 
We describe complementary insights from clustering by direct citation and co-citation. \deleted[remark=repositioned]{, and both point to the increase in computer science publications and their scope}.
Our analysis \replaced{reveals}{shows} cross-category clusters, some that interact with external fields, such as the biological sciences, while others remain inward looking.\added{Overall, we document an increase in computer science publications and their scope.}

@@

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. As observed in a 2017 US National Academies Report, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission"~\cite{nas_2017}. More recently, the collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have been \replaced{presented as a driver for digital transformation}{postulated as a recent driver for growth and evolution}~\cite{siebel2019_digital}. Given this \replaced{rapid growth and expansion}{powerful influence}, an updated understanding of the present state and structure of computer science and its relationship to other fields \replaced{can only inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy.}{could inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy.}

@@

Other classification systems are available, such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the National Science Foundation classification system~\cite{nsf_classification,scopus_ref,wos_ref}. Scopus and the Web of Science also include comprehensive bibliographies with citation links and proprietary unique identifiers for publications. The Scopus All Science Journal Classification (ASJC), which we use in this study, is organized  into 4 subject areas, 27 major subject areas, and 334 minor subject areas. All three systems rely on applying one or more journal-derived labels to articles.  A logical prediction, given the diversity of articles within journals, is limited specificity at the article level. 
Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD)  and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Others have also discussed and critiqued disciplinary assignments using journal-based classifications~\cite{wang_large-scale_2016,perianes-rodriguez_comparison_2017}. Article classification systems have been constructed that escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, but do not seem to presently enjoy widespread use.

\replaced{The purpose of this study is to develop an improved understanding of the structure of the field of computer science relative to the landmark study of Salton and Bergmark forty years ago. We extend their work by takimg advantage of modern bibliographic resources and clustering technologies and using a combination of article and journal approaches to study trends in the computer science literature. We also consider connections to other fields, especially biology, since approximately 42\% of Scopus is classified under the top level subject areas of Life Sciences and Health Sciences. As a source of computer  literature, we use DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research.} {In this study, we extend the work of Salton and Bergmark through a combination of article and journal approaches to study trends in computer science, while also considering connections to other fields, especially biology since approximately 42\% of Scopus is classified under the top level subject areas of Life Sciences and Health Sciences. As a source of computer  literature, we used DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research. We describe below the high-level interactions of computer science within the Physical Sciences, and with the Social Sciences, Life Sciences, and Health Sciences.}

\added{We constructed clusters of articles using either direct citation or co-citations and match these clusters to ASJC categories to place them in context. We also align clusters from direct citations to those from co-citations and discuss our findings, Overall, we note cross category clusters with different degrees of interaction with fields outside computer science.}

@@

\emph{Overview} \added[id=gc,{remark=Reviewer 1}]{The underlying assumption is that, notwithstanding incomplete coverage, records in DBLP are greatly enriched for computer science~\cite{dblp_ref}. Therefore, the contents of DBLP are an excellent sample of the computer science literature to analyze. The workflow described in the following paragraphs of Materials and Methods, consists at a high-level of (i) merging the DBLP and Scopus datasets using digital object identifiers(DOIs) (ii) extracting cited references for the articles common to DBLP and Scopus (iii) clustering by direct citation and by our modification of co-citation based clustering. The outcome was a dataset of 8,000,411 publications.}

\replaced{Our working definition of the computer science literature, for the purpose of this study, was all publications in the DBLP bibliography that (i) had a digital object identifier DOI, and (ii) could also be matched to article identifiers in the Scopus bibliography.}{For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography.} Cross-matching DBLP publications to records in the Scopus abstract and citation database of peer-reviewed literature enables us to harvest the richer links in Scopus, as well as \added{to} extract links to publications from other disciplines. Cross-matching to Scopus also allows the use of journal-\added{based} classifications when clustering documents at the article-level. We used DBLP articles from journals and conference proceedings to construct article clusters by using either direct citations or co-citations as links. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications, which when combined with cited references extracted from Scopus grew to 8,000,411 publications.

\added{\emph{DBLP data}}A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Slightly over 95\% of the publications within were published after 1996. Publications were parsed from the XML source file and loaded into a PostgreSQL database.

\added{\emph{Scopus data}} As part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}, we have previously parsed the Scopus dataset, presently at over 88 million publications, into a custom schema in a PostgreSQL database. 
The total number of publications in Scopus labeled with major subject area Computer Science (in turn a subset of the Physical Sciences subject area) is 5,835,160.

\added{Merging and graph construction} Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings.  References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 publications and references.

\added{We represented these  8,000,411 records, referred to as the  \emph{comp} dataset (Fig.~\ref{fig:ar_cp_annotation}, Table~\ref{tab:comp}), as a graph where the 8,000,411 nodes represent publications and references  and the 44,296,381 undirected edges represent citations within the dataset.}

\emph{Clustering} Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation, with direct citation being proposed as as the best approach to concentrate citation links~\cite{kessler_comparison_1965,klavans_which_2017}. Accordingly, we used direct citation links as the basis for cluster formation, and also co-citation to obtain an alternative view. In applying both clustering by direct citation and by co-citation, we attempted to consider, wherever possible, the criteria articulated by \v{S}ubelj, van Eck, and Waltman~\cite{subelj_clustering_2016} that (i) the largest cluster should be no more than 10 times the smallest one, (ii) small clusters should be eliminated, (iii) small changes and replicates should yield similar results (``stability"),  (iv) computing time should be minimized where possible, and (v) the clustering should seem reasonable on a qualitative level (``intuitive sensibility").

 \emph {Direct Citation} Graclus~\cite{graclus_2007} is a spectral graph clustering package that optimizes  various clustering criteria, including normalized cut, ratio cut, and ratio association, and that has previously been applied to citation data~\cite{subelj_clustering_2016}. 
We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At around 20 clusters, clusters size was relatively stable with the largest cluster containing roughly 10 times the number of nodes in the smallest one,
so that 20 clusters is a good choice with respect to the criteria specified in~\cite{subelj_clustering_2016}. Consequently, we used Graclus to generate 20 clusters, labeled 0-19 (Table~\ref{tab:graclus}), analogous to Level 1 of Waltman and van Eck's mapping of nearly 10 million publications but focused on the DBLP bibliography rather than a broader Web of Science sample~\cite{waltman_new_2012}.

We also used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering by direct citation (smaller is better), noting that conductance has been found to be a good metric for this purpose~\cite{emmons2016analysis,almeida_2012}. In our analysis, we saw that the last cluster (cluster 19) had a much larger conductance value than the other clusters and also had the smallest number of nodes. We then examined results obtained using Graclus with two other numbers of clusters (18 and 22), and in each case, the highest numbered cluster had the greatest conductance value and also the smallest number of nodes. These results suggest that Graclus produces a final cluster that effectively serves as a container for `left over publications' during the clustering procedure. Therefore, we limited our consideration of cluster 19 (the final of the twenty clusters)  when interpreting results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15 (Fig.~\ref{fig:graclus_comparison_conductance}, Table~\ref{tab:comp}).

@@

\added{We began by identifying highly cited articles in \emph{comp} and selected those in the 90th percentile (212,311 articles). We then identified ~4.3 million publications in Scopus that cite these 212,311 articles. For each of the 4.3 million citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs generated were then restricted to those where both members of a pair were in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979} to limit dominance by areas with high citation activity. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par }

@@

\deleted{\emph{Data}The process of selection and matching resulted in a dataset of  publications (Materials and Methods). Of 4,291,130 DBLP publications with DOIs, only  2,685,356 had corresponding DOIs in Scopus.} Of 2.68 million publications in \emph{comp}, approximately 2.07 million were assigned ASJC codes in Scopus corresponding to Computer Science (major subject area with 13 minor subject areas (Table~\ref{tab:comp}), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (Fig.~\ref{fig:ar_cp_annotation}). The set of 2.07 million publications classified under the major subject area Computer Science in Scopus spanned all 13 minor subject areas with Software at 30.3\%  being the largest component and Computer Science (miscellaneous) at 0.9\% the smallest. Publications in the \emph{comp} dataset labeled Theoretical Computer Science (409,082) are classified under the the major subject area of Mathematics rather than Computer Science (Table~\ref{tab:comp}).

\iffalse
Overall, 38.86\% of the publications were assigned to one minor subject area, 78.1\% to two minor subject areas, 90.9\% to three minor subject areas, and 0.0003\% assigned to 8 different minor subject areas. 
\fi

\emph{Clustering by Direct Citation} We constructed article-level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to the Scopus ASJC classification. To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster.

@@

\section{Discussion} 
\added{Considering expansion and diversification of the field of computer science, we revisited its characterization by Salton and Bergmark in 1979~\cite{salton_citation_1979}. In comparison to \cite{salton_citation_1979}, we analyzed considerably more data, 2.68 million publications versus 391 by using two bibliographic databases, DBLP and Scopus, consisting of  approximately 7 million and 88 million publications respectively. By linking the two datasets, we were able to harvest citation data as well as other metadata that enabled theo construct article level clusters in two different ways and reconciliation with the Scopus ASJC classification, for a journal level perspective.}

@@

The focus of this article was on high-level features, and a preference for simplicity and intuitiveness in the choice of methods. This preference influences the specific choices we made, such as the (i) use of the DBLP dataset matched to Scopus, which may not capture all aspects of computer science and its interactions with other fields, (ii) the basis for clustering, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. We believe that an approach that combines journal-level  with article-level analyses is useful for studies of this kind. \added{However, we acknowledge that our approach would not detect emerging fields with few publications that could be important signs of innovation. We also feel strongly that contextualized interpretation by domain experts should be performed on results generated by such approaches.}

We speculate that biology, often dominant in bibliometric studies, is restricted to two clusters on account of (i) the focus of DBLP,  (ii) our use of normalized co-citations, and (iii) the threshold set for detection. Future research should, of course, include complementary investigations at finer levels of granularity and sensitivity using article-level and topic approaches that others have developed~\cite{glanzel_using_2017,boyack_clustering_2011,boyack_investigating_2017,boyack_classification_2014,sjogarde_granularity_2019,traag_louvain_2019}. We also refer readers to a related study of the computer science literature~\cite{chakraborty_role_2018} that is focused on evolving interdisciplinarity in computer science using data from Microsoft Academic Research and a classification of Computer Science into 24 categories.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.85\textwidth]{quad_plot.pdf}
%\includegraphics[width=0.9\textwidth]{quad-chart-narrow.pdf}
\caption{Co-citation analysis.
(a) Schematic representation of variable clustering protocol modified from Small and Sweeney (1985) \cite{small_clustering_1985}. Three parameters are specified: (i) a threshold or starting level based on a quantile of normalized co-citation frequency, (ii) a level increment, and  (iii) a maximum cluster size. Input data is a set of co-cited publications with edge-weight defined by normalized co-citation frequencies. Green clusters are within the max cluster size. At the initial threshold, $t1$, a single cluster below the maximum cluster size, $mcs$ (green), along with one large cluster above it (red) are generated. As the threshold is incremented to $t2$, additional clusters of acceptable size is generated. The cascade continues to completion, which is defined by all clusters being of size less than or equal to the $mcs$. In this schematic, five rounds are adequate for the process to run to completion.
(b) The distribution of publications (using fractional counting) across four top-level ASJC subject areas after applying variable level clustering as in a)
(c) The Venn diagram of the fractional counting given in (b).
(d) The distribution of cluster sizes (logarithmic y-axis) as a function of the number of iterations of the agglomerative clustering technique; note that the largest cluster is extremely large when the number of iterations exceeds 600.
}
\label{fig:quad-chart}
\end{figure}

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.6]{graclus_cocit_fig.pdf}
% figure caption is below the figure
\caption{Intersections between clusters generated using direct citation and co-citation features. x-axis: Clusters generated by Graclus number 0-19. y-axis: Cluster generated by variable level clustering and agglomerative clustering using a modification of Small and Sweeney (1985)~\cite{small_clustering_1985}. Point size (perc) is the percentage of a co-cited cluster that maps to a corresponding Graclus cluster. A minimum threshold of 15\% was set. Graclus cluster 19, the 20th cluster, did not map to any cluster on the y:axis.}
\label{fig:graclus_cocit_fig}       % Give a unique label
\end{figure}
\newpage
\clearpage
\section{Tables}

@@

\begin{acknowledgements} The authors thank Henry Small for very helpful discussions. Research and development reported in this publication was partially supported by funds from the National Institute on Drug Abuse, National Institutes of Health, US Department of Health and Human Services, under Contract No HHSN271201800040C (N44DA-18-1216). TW is supported by the Grainger Foundation. Citation data used in this paper relied on Scopus (Elsevier Inc.) as implemented in the ERNIE project (Korobskiy et al., 2019), which is collaborative between NET ESolutions Corporation and Elsevier Inc. We thank our Elsevier colleagues for their support of the ERNIE project.
\end{acknowledgements}

>>>>

\usepackage{rotating}
\usepackage{pdflscape}

@@

\begin{abstract}
Computer science has experienced dramatic growth and diversification over the last twenty years. Towards a current understanding of the structure of this discipline, we analyze 
a \replaced{large sample}{cohort} of the computer science literature from the DBLP database. For insight on the features of this cohort and the relationship within its components, we have constructed article 
level clusters based on either direct citations or co-citations, and reconciled them \replaced{with}{to} major and minor subject categories in the All Science Journal Classification (ASJC). 
We describe complementary insights from clustering by direct citation and co-citation, and both point to the increase in computer science publications and their scope.
Our analysis \replaced{reveals}{shows} cross-category clusters, some that interact with external fields, such as the biological sciences, while others remain inward looking. \added{Overall, we document an increase in computer science publications and their scope.}

@@

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. As observed in a 2017 US National Academies Report, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission"~\cite{nas_2017}. More recently, the collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have been presented as a driver for digital transformation~\cite{siebel2019_digital}. Given this rapid growth and expansion, an updated understanding of the present state and structure of computer science and its relationship to other fields can inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy.

@@

Other classification systems are available, such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the National Science Foundation classification system respectively~\cite{nsf_classification,scopus_ref,wos_ref}. Scopus and the Web of Science are commercially available databases with comprehensive, though not identical coverage~\cite{archambault_2009}, that can be used to conduct large scale analyses. The All Science Journal Classification (ASJC), maintained by Scopus, which we use in this study, is organized  into 4 subject areas, 27 major subject areas, and 334 minor subject areas. All three systems rely on applying one or more journal-derived labels to articles.

A logical prediction, given the diversity of articles within journals, is limited specificity at the article level. Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD)  and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Others have also discussed and critiqued disciplinary assignments using journal-based classifications~\cite{wang_large-scale_2016,perianes-rodriguez_comparison_2017}. Article classification systems have been constructed that escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, but do not seem to presently enjoy widespread use.

The purpose of this study is to develop an improved understanding of the structure of the field of computer science relative to the landmark study of Salton and Bergmark forty years ago. We extend their work by taking advantage of modern bibliographic resources and clustering technologies and using a combination of article and journal approaches to study trends in the computer science literature. We also consider connections to other fields, especially biology, since approximately 42\% of Scopus is classified under the top level subject areas of Life Sciences and Health Sciences. As a source of computer  literature, we use DBLP, a reference bibliography for computer science~\cite{dblp_ref,pham_2010}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research. We construct clusters of articles using either direct citation or co-citations and match these clusters to ASJC categories to place them in context. We also aligned clusters from direct citations to those from co-citations and discuss our findings. Overall, we note cross category clusters with different degrees of interaction with fields outside computer science.

@@

\emph{Overview} Our underlying assumption is that, notwithstanding incomplete coverage, records in DBLP are greatly enriched for computer science~\cite{dblp_ref,pham_2010}. Therefore, the contents of DBLP are an excellent sample of the computer science literature to analyze. The workflow described in the following paragraphs of Materials and Methods, consists at a high-level of (i) merging the DBLP and Scopus datasets using digital object identifiers(DOIs) (ii) extracting cited references for the articles common to DBLP and Scopus (iii) clustering by direct citation and by our modification of co-citation based clustering. The outcome was a dataset of 8,000,411 publications.

Our working definition of the computer science literature, for the purpose of this study, was all publications in the DBLP bibliography that (i) had a digital object identifier DOI, and (ii) could also be matched to article identifiers in the Scopus bibliography. Cross-matching DBLP publications to records in the Scopus abstract and citation database of peer-reviewed literature enables us to harvest the richer links in Scopus, as well as extract links to publications from other disciplines. Cross-matching to Scopus also allows the use of journal-\added{based} classifications when clustering documents at the article-level. We used DBLP articles from journals and conference proceedings to construct article clusters using either direct citations or co-citations as links. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications, which when combined with cited references extracted from Scopus grew to 8,000,411 publications.

\emph{DBLP data.} A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Slightly over 95\% of the publications within were published after 1996. Publications were parsed from the XML source file and loaded into a PostgreSQL database.

\emph{Scopus data.} As part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}, we have previously parsed the Scopus dataset, presently at over 88 million publications, into a custom schema in a PostgreSQL database. The total number of publications in Scopus labeled with major subject area Computer Science (in turn a subset of the Physical Sciences subject area) is 5,835,160.

\emph{Merging and graph construction.} Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings.  References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 publications and references.

We represented these  8,000,411 records, referred to as the  \emph{comp} dataset (Fig.~\ref{fig:ar_cp_annotation}, Table~\ref{tab:comp}), as a graph where the 8,000,411 nodes represent publications and references  and the 44,296,381 undirected edges represent citations within the dataset.

\emph{Clustering.} Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation, with direct citation being proposed as as the best approach to concentrate citation links~\cite{kessler_comparison_1965,klavans_which_2017}. Accordingly, we used direct citation links as the basis for cluster formation, and also co-citation to obtain an alternative view. In applying both clustering by direct citation and by co-citation, we attempted to consider, wherever possible, the criteria articulated by \v{S}ubelj, van Eck, and Waltman~\cite{subelj_clustering_2016} that (i) the largest cluster should be no more than 10 times the smallest one, (ii) small clusters should be eliminated, (iii) small changes and replicates should yield similar results (``stability"),  (iv) computing time should be minimized where possible, and (v) the clustering should seem reasonable on a qualitative level (``intuitive sensibility").

\emph {Direct Citation.} Graclus~\cite{graclus_2007} is a spectral graph clustering package that optimizes  various clustering criteria, including normalized cut, ratio cut, and ratio association, and that has previously been applied to citation data~\cite{subelj_clustering_2016}. We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At around 20 clusters, clusters size was relatively stable with the largest cluster containing roughly 10 times the number of nodes in the smallest one,
so that 20 clusters is a good choice with respect to the criteria specified in~\cite{subelj_clustering_2016}. Consequently, we used Graclus to generate 20 clusters, labeled 0-19 (Table~\ref{tab:graclus}), analogous to Level 1 of Waltman and van Eck's mapping of nearly 10 million publications but focused on the DBLP bibliography rather than a broader Web of Science sample~\cite{waltman_new_2012}.

We also used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering by direct citation (smaller is better), noting that conductance has been found to be a good metric for this purpose~\cite{emmons2016analysis,almeida_2012}. In our analysis, we saw that the last cluster (cluster 19) had a much larger conductance value than the other clusters and also had the smallest number of nodes. We then examined results obtained using Graclus with two other numbers of clusters (18 and 22), and in each case, the highest numbered cluster had the greatest conductance value and also the smallest number of nodes. These results suggest that Graclus produces a final cluster that effectively serves as a container for `left over publications' during the clustering procedure. Therefore, we limited our consideration of cluster 19 (the last of the twenty clusters)  when interpreting our results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15 (Fig.~\ref{fig:graclus_comparison_conductance}, Table~\ref{tab:comp}).

@@

We began by identifying highly cited articles in \emph{comp} and selected those in the 90th percentile (212,311 articles). We then identified ~4.3 million publications in Scopus that cite these 212,311 articles. For each of the 4.3 million citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs generated were then restricted to those where both members of a pair were in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979} to limit dominance by areas with high citation activity. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par

@@

Of 2.68 million publications in \emph{comp}, approximately 2.07 million were assigned ASJC codes in Scopus corresponding to Computer Science (major subject area with 13 minor subject areas (Table~\ref{tab:comp}), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (Fig.~\ref{fig:ar_cp_annotation}). The set of 2.07 million publications classified under the major subject area Computer Science in Scopus spanned all 13 minor subject areas with Software at 30.3\%  being the largest component and Computer Science (miscellaneous) at 0.9\% the smallest. Publications in the \emph{comp} dataset labeled Theoretical Computer Science (409,082) are classified under the the major subject area of Mathematics rather than Computer Science (Table~\ref{tab:comp}).

\emph{Clustering by Direct Citation.} We constructed article-level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to the Scopus ASJC classification. To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster.

@@

\section{Discussion} 
Considering expansion and diversification of the field of computer science, we revisited its characterization by Salton and Bergmark in 1979~\cite{salton_citation_1979}. In comparison to \cite{salton_citation_1979}, we analyzed considerably more data, 2.68 million publications versus 391 by using two bibliographic databases, DBLP and Scopus, consisting of  approximately 7 million and 88 million publications respectively. By linking the two datasets, we were able to harvest citation data as well as other metadata that enabled theo construct article level clusters in two different ways and reconciliation with the Scopus ASJC classification, for a journal level perspective.

@@

The focus of this article was on high-level features, and a preference for simplicity and intuitiveness in the choice of methods. This preference influences the specific choices we made, such as the (i) use of the DBLP dataset matched to Scopus, which may not capture all aspects of computer science and its interactions with other fields, (ii) the basis for clustering, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. We believe that an approach that combines journal-level  with article-level analyses is useful for studies of this kind. However, we acknowledge that our approach would not detect emerging fields with few publications that could be important signs of innovation. We also feel strongly that contextualized interpretation by domain experts should be performed on results generated by such approaches.

We speculate that biology, often dominant in bibliometric studies, is restricted to two clusters on account of (i) the focus of DBLP,  (ii) our use of normalized co-citations, and (iii) the threshold set for detection. Future research should, of course, include complementary investigations at finer levels of granularity and sensitivity using article-level and topic approaches that others have developed~\cite{glanzel_using_2017,boyack_clustering_2011,boyack_investigating_2017,boyack_classification_2014,sjogarde_granularity_2019,traag_louvain_2019}. We also refer readers to related studies of the computer science literature~\cite{chakraborty_role_2018} that are focused on (i) evolving interdisciplinarity in computer science using data from Microsoft Academic Research and a classification of Computer Science into 24 categories (ii) combining the DBLP and CiteSeerX libraries to generate a knowledge network~\cite{pham_2010}.

@@

\newpage
\begin{landscape}
\begin{figure}[ht]
\centering
\includegraphics[width=1.4\textwidth]{vlc_thrice.pdf}
%\includegraphics[width=0.9\textwidth]{quad_plot.pdf}
%\includegraphics[width=0.9\textwidth]{quad-chart-narrow.pdf}
\caption{Co-citation analysis.
(a) Schematic representation of variable clustering protocol modified from Small and Sweeney (1985) \cite{small_clustering_1985}. Three parameters are specified: (i) a threshold or starting level based on a quantile of normalized co-citation frequency, (ii) a level increment, and  (iii) a maximum cluster size. Input data is a set of co-cited publications with edge-weight defined by normalized co-citation frequencies. Green clusters are within the max cluster size. At the initial threshold, $t1$, a single cluster below the maximum cluster size, $mcs$ (green), along with one large cluster above it (red) are generated. As the threshold is incremented to $t2$, additional clusters of acceptable size is generated. The cascade continues to completion, which is defined by all clusters being of size less than or equal to the $mcs$. In this schematic, five rounds are adequate for the process to run to completion.
(b) The distribution of publications (using fractional counting) across four top-level ASJC subject areas after applying variable level clustering as in a)
(c) The Venn diagram of the fractional counting given in (b).
(d) The distribution of cluster sizes (logarithmic y-axis) as a function of the number of iterations of the agglomerative clustering technique; note that the largest cluster is extremely large when the number of iterations exceeds 600.
}
\label{fig:quad-chart}
\end{figure}
\end{landscape}
\newpage

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.6]{graclus_cocit_fig.pdf}
% figure caption is below the figure
\caption{Intersections between clusters generated using direct citation and co-citation features. x-axis: Clusters generated by Graclus number 0-19. y-axis: Cluster generated by variable level clustering and agglomerative clustering using a modification of Small and Sweeney (1985)~\cite{small_clustering_1985}. Point size (perc) is the percentage of a co-cited cluster that maps to a corresponding Graclus cluster. A minimum threshold of 15\% was set. Graclus cluster 19, the 20th cluster, did not map to any cluster on the y-axis.}
\label{fig:graclus_cocit_fig}       % Give a unique label
\end{figure}
\newpage
\clearpage
\section{Tables}

@@

\begin{acknowledgements} The authors thank Henry Small for very helpful discussions. Research and development reported in this publication was partially supported by funds from the National Institute on Drug Abuse, National Institutes of Health, US Department of Health and Human Services, under Contract No HHSN271201800040C (N44DA-18-1216). TW is supported by the Grainger Foundation. Citation data used in this paper relied on Scopus data as implemented in the ERNIE project (Korobskiy et al., 2019), which is collaborative between NET ESolutions Corporation and Elsevier Inc. We thank our Elsevier colleagues for their support of the ERNIE project.
\end{acknowledgements}

% Authors must disclose all relationships or interests that 
% could have direct or potential influence or impart bias on 
% the work: 
%
 \section*{Conflict of interest}
The authors declare that they have no conflicts of interest. Elsevier personnel played no role in conceptualization, experimental design, review of results, or conclusions presented. The content of this publication is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health, NET ESolutions Corporation, or Elsevier Inc.
